---
title: "Homework 4"
author: "Tiana Walker"
date: "9/20/21"
output: pdf_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

Write a brief description in your own words of each of the types of chart or graph below: 

### (a) Bar Chart
a bar chart is like a histogram in that it represents the distribution of data in vertical bars that correspond with the y-axis. However, the x-axis usually isn’t numerical; rather it is data that is proportional to the values or info they represent. 

### (b) Scatter Plot
a scatter plot is a graph where there are multiple variables spread out across the x and y axes to show their relation to each other based on the data provided
### (c) Line Plot
- a line plot is a graph where there are multiple variables spread out across the x and y axes, but the points are connected by a line to show the overall spread/frequency of the data. 
### (d) Box Plot
a box plot is a way to depict groups of data by five different categories: the minimun, maximum, sample media, and first and third quartiles. 

### (e) Histogram
a histogram represents a visualization of an approximate numerical distribution of data via vertical bars placed along the x axis, with the vertical bars showing its place along the y axis. 


# Problem 2
The following problems ask you to analyze some `bad' visualizations using the frameworks from class. 

### (a) 
Pick one of the `bad' visualizations on http://www.perceptualedge.com/examples.php  (click on the images to see the descriptions) and summarize in your own words the changes that were made to update the plot. 

I chose the graph of the Sept 2003 issue DM Review. The original pie chart is extremely messy and cluttered and very difficult to understand. The sectioning isn't clear and has no coloring to tell the reader which part of the chart the info corresponds to. Instead, the modified version presents the data much clearer. They use a horizontal bar chart to visualize the data instead, and it makes much more sense and is tidy and easy to read and comprehend.

### (b)
Select one of the bad visualizations on  https://www.businessinsider.com/the-27-worst-charts-of-all-time-2013-6  and provide brief responses to  the following questions: 

* (i) What is the visualization trying to communicate?
I chose the unemployment rate chart by Fox news. I think they are trying to communicate that unemployment rates are high, or that there should be cause for concern because there is so much red area? It is honestly hard to tell what they are even trying to communicate because the graph is so embarrassingly poorly done. 
* (ii) Why does it fail to be effective?
I chose the unemployment rate chart by Fox news. I think they are trying to communicate that unemployment rates are high, or that there should be cause for concern because there is so much red area? It is honestly hard to tell what they are even trying to communicate because the graph is so embarrassingly poorly done.
* (iii) How would you modify the plot to improve it according to the principles we discussed in class?  
I would make the graph's y axis start at 0 and go to 100, and use the line plot to show that there is hardly and change or difference in a year at all. 

# Problem 3

Load the data in GME_Stock.csv into R as a data frame or tibble. For parts (a), (b), and (c) below, create the plots using this dataset, making sure to label the axes and follow the other principles of good figure design: 
```{r gme}
gme <- read.csv("GME_stock.csv")

head(gme)
library("tidyverse")
head(gme)
ecd <- economics 
```


### (a) 
Plot the values in the closing prices column as a line 
```{r plot}
plot(gme$Close,type="l",ylab="Closing Prices", col= "green")
```



### (b)
Make a scatterplot of opening prices vs. closing prices.
```{r plot2}
plot(gme$Open ~ gme$Close, xlab="Closing Prices", ylab="Opening Prices", main= "Opening vs. Closing")
```

### (c)
Add a new column to your dataframe whose rows represent the change in price during each day and make a histogram of these values. 
```{r pricechange}
gme$Price_Change <- (gme$Close- gme$Open)
```
```{r plot9}
hist(gme$Price_Change, main = "Change in Price")
```

### (d)
Make a bar chart presenting the in-state tuition data presented in this table:


\begin{tabular}{|l|c|c|c|c|c|}
\hline
School&WSU&EWU&UW&UI&CWU\\
\hline
Tuition&11841& 7526&11465&8304&8273\\
\hline
\end{tabular}\\

```{r tuition}
tuition= data.frame(
   schools = c("EWU", "CWU", "UI", "UW", "WSU"),
   tuitions =c(7526,8273,8304,11465,11841)
)
```

```{r plot4}
ggplot(data=tuition, mapping=aes(x=schools, y=tuitions)) + geom_col()
```
### (e)
Provide a brief justification for the order you placed the bars from left to right in (d).
Well, I tried to place them in ascending order but I don't know why it didn't output in ascending order. But this barchart basically displays the amount of tuition on the y axis and the x axis displays each school. 

# Problem 4

Read the Readme for the MovieRatings Dataset and pick one of the sub tables (such as  byyearbyage or disneymovies) that interest you the most to do some exploratory analysis. 

### (a) 
 Load your chosen table as a dataframe or tibble in R.
```{r movies}
movies <- read.csv("MovieRating_avgRatingByMovieByGender.csv")
```

### (b)
Explore the data by making at least two of the charts or plots we discussed this week. 
```{r plot5}
plot(movies$avg.age.women ~ movies$avg..age.men, xlab="Average Age of Men", ylab="Average Age of Women", main= "Women vs. Men")
```
```{r plot6}
ggplot(data=movies, mapping = aes(x= avg.age.women, y= avg.rating.women, color= avg.age.women)) +geom_point(size=1) 
``` 
```{r agedif}
movies$Age_Dif <- (movies$avg.age.women- movies$avg..age.men)
```
```{r plot7}
hist(movies$Age_Dif, main = "Age Difference")
```
```{r plot8}
barplot(table(movies$avg.age.women))
```


### (c)
 Select one pattern that is revealed by your visualizations.
### (d)
For my scatterplot I compared the average rating of women and the average age of women. One pattern that I noticed is that most of the ratings were submitted by women under the age of 50. 

Write a paragraph describing the data, the pattern that you observed, and how your plot displays that pattern.

So on the left/the y axis is the average rating of women. On the bottom/x axis, there is the average age of women. I added the color scale to help it make more sense; otherwise it would have just been pretty much a big black blob with a few outliers. I noticed that women the age of 50 and older hardly wrote reviews at all, and they hardely ever wrote bad/1 star reviews. Most of the bad/1-3 star reviews are from younger women, below the age of thirty. The majority of the high 4-5 star reviews are from women between the age of 30-45. You can tell this from the color disbursement across the x and y axes. 
### (e)

Come up with a potential explanation for the trend that you described in (d) and write a couple of sentences describing the what data you could use to test you explanation further. 

I'm guessing that the reason older women hardly write reviews is because they are on the internet a lot less, and aren't as tech savvy or tuned into pop culture. They might not even know that they can write reviews. I really don't know why younger women tended to give worse reviews. Maybe its because they have seen more movies and have higher expectations? I don't know. I think something we could use to test the data further is to also compare men's rating vs. age trends, and see if they are similar. If they are, then we could probably confirm that older folks probably in general aren't leaving reviews. We could also see if there is a trend with men that shows that younger men give worse reveiws too. If that is true, then we can just probably assume it is because of a generational gap. If the comparisons don't match up, then maybe we'd just need to go back with another analysis and ask different questions. 
